---
title: "Jewitt_Ally_A9"
author: "Ally Jewitt"
date: "2023-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r}
library(doParallel)
library(parallel)
library(foreach)
```

## Number of cores
```{r}
detectCores()
```

### There are 4 cores in my system. 

## Standard for loop
```{r}
mean <- numeric(4000)

start1 <- Sys.time()

for (i in 1:4000) {
  sample <- rnorm(100000, mean=10, sd=3)
  mean[i] <- mean(sample)
}

finish1 <- Sys.time()

time_serial <- finish1 - start1
print(paste0("Time elapsed: ", time_serial, " seconds"))
```

## Theoretical time 

Theoretically, the parallel run-time should be equivalent to the serial run-time, divided by the number of cores in my computer:
```{r}
SerialTime<-13.4488620758057
Comp_Core<- 4

time_theoretical<-SerialTime/Comp_Core
print(paste0("Theoretical time: ", time_theoretical, " seconds"))
```


## Parallel loop 
```{r}
Cores <- parallel::makeCluster(detectCores()) 
doParallel::registerDoParallel(Cores) 

start2 <- Sys.time()

means<- foreach(i = 1:4000, .combine = c) %dopar% {
  numbers <- rnorm(100000, mean = 10, sd = 3)
  mean(numbers)
}

parallel::stopCluster(Cores) 

finish2 <- Sys.time()

time_parallel <- finish2 - start2
print(paste0("Time elapsed: ", time_parallel, " seconds"))
```

## Compare run times 
The actual run time of the parallel loop is 13.7 seconds, which is faster than the serial loop, 30.4 seconds, but slower than the theoretical, 3.4 seconds. 

The theoretical is shorter than the actual run-time of the parallel program because it represents the minimum amount of time that the system needs to run the loop. This is often a bit slower in actual time because of other factors such as contention for resources. 



